\section{Discrete Probability Distributions}

\setcounter{subsection}{1}
\subsection{Indicator Variables}

\begin{enumerate}
    \item \textbf{TBD}
    \item \textbf{A special case of Boole's inequality occurs when the events all have the same probability. 
    Suppose $n$ graduates all throw their caps in the air and then retrieve a cap at random. 
    Find an expression for the probability that none of the students retrieve their own cap (a derangement). 
    Find the limit of this probability if the number of caps $n \rightarrow \infty$. \textit{Hint: For $i = 1, . . . , n,$ let $A_{i}$ represent the event that person $i$ retrieves their own cap. 
    Then $(A_1 \cup A_{2} \cup . . . \cup A_{n})^{c}$ is the event that nobody ends up with their own cap.}} \\
    The probability of a derangement is
    \[
        P(A_{1} \cup A_{2} \cup . . . \cup A_{n})^{c} = 1 - P(A_{1} \cup A_{2} \cup . . . \cup A_{n}).\] 
    Taking the limit of this expression as $n \rightarrow \infty$ and using the Principle of Inclusion Exclusion, we know that this
    \begin{align*}
    \lim\limits_{n \rightarrow \infty} 1 - P(A_{1} \cup A_{2} \cup . . . \cup A_{n}) &= 1 - \left[\sum\limits^{n}_{k=1}P(A_{i}) - \sum\limits_{i<j}P(A_{i} \cap A_{j}) + \sum\limits_{i<j<k}P(A_{i} \cap A_{j} \cap A_{k}) - \dots\right] \\
    &= \lim\limits_{n \rightarrow \infty} 1 - \left[n \cdot \frac{1}{n} - \binom{n}{2} \left(\frac{1}{n} \cdot \frac{1}{n-1} \right) + \binom{n}{3} \left(\frac{1}{n} \cdot \frac{1}{n-1} \cdot \frac{1}{n-2} \right)  \right]
    \end{align*}
    Simplifying, this becomes
    \[
        \lim\limits_{n \rightarrow \infty}1- \left[ 1 - \frac{1}{2!} + \frac{1}{3!} - \frac{1}{4!} + \dots \right] = \lim\limits_{n \rightarrow \infty} \frac{1}{2!} - \frac{1}{3!} + \frac{1}{4!} - \dots
    \]
    
    Recall the Taylor Expansion of $e^{x}$: $e^{x} = \sum\limits^{\infty}_{k=0} \frac{x^{k}}{k!}.$ 
    We recognize the above expression as $e^{-1}$: 
    \begin{align*}
        e^{-1} = \sum\limits^{\infty}_{k=0} \frac{(-1)^{k}}{k!} &= \lim\limits_{n \to \infty} \left(1-1 + \frac{1}{2!} - \frac{1}{3!} + \frac{1}{4!} - \dots \right) \\
        &= \lim\limits_{n \to \infty} \frac{1}{2!} -\frac{1}{3!} + \frac{1}{4!} - \dots 
    \end{align*}
    Thus,
    \[ 
        \lim\limits_{n \to \infty} P(A_{1} \cup A_{2} \cup . . . \cup A_{n})^{c} = \boxed{\frac{1}{e}}.
    \]
        
    \item \textbf{In the caps example, let $X$ represent the number of graduates who retrieve their own cap. Explain why the distribution of $X$ is approximately
    Poisson$(1)$ when $n$ is large (see 1(c)). Compare the exact probabilities of $X = 0$ and $X=1$ to the corresponding Poisson probabilities when $n=5$.}

    Note that the probability that a given graduate receives their own cap (assuming nothing about all other graduate cap arrangements) is $p = \frac{1}{n}$. Though we are essentially sampling without replacement when throwing all the graduate caps in the air and retrieving them at random, we know from problem 1(b) that the distribution $X$ for the number of graduates who receive their own cap converges to $\mathrm{Binom}(n, p)$ when $n$ is large. \\

    Furthermore, note that $\lambda = np = n\frac{1}{n} = 1$ is fixed. Consequently, the limit of $P(X = x)$ as $n$ is large, by problem 1(c), is simply $\mathrm{Poisson}(\lambda) = \mathrm{Poisson}(1)$. We conclude that the distribution of $X$ is approximately $\mathrm{Poisson}(1)$ when $n$ is large. \\

    Using the approximation $X \sim \mathrm{Poisson}(1)$ for large $n$, we can approximate the probability of a derangement
    \[ 
        f_x(x=0, \lambda = 1) = \frac{1^0 e^{-1}}{0!} = \frac{1}{e}
    \]
    and the probability that exactly one graduate gets their own cap:
    \[ 
        f_x(x=1, \lambda = 1) = \frac{1^1 e^{-1}}{1!} = \frac{1}{e}.
    \]

    We can also compare these approximations with the exact probabilities of $X=0$ and $X=1$ for $n=5$. Note that the probability of a derangement for $n=5$ can be calculated using the formula from 2(b):
    \begin{align*}
        P(X = 0)&= 1 - \left[ 5 \cdot \frac{1}{5} - \binom{5}{2}\frac{1}{5} \cdot \frac{1}{4} + 
        \binom{5}{3}\frac{1}{5} \cdot \frac{1}{4} \cdot \frac{1}{3} - \binom{5}{4}\frac{1}{5} \cdot \frac{1}{4} \cdot \frac{1}{3} \cdot \frac{1}{2}
        + \binom{5}{5}\frac{1}{5} \cdot \frac{1}{4} \cdot \frac{1}{3} \cdot \frac{1}{2} \cdot \frac{1}{1}\right] \\
        &= \frac{11}{30} \approx 0.3\overline{6}
    \end{align*}

    Note that the probability that $X=1$ can be calculated by ``picking'' the graduate to get their own cap (of which there are $5$ possibilites),
    multiplying this by the probability that the chosen graduate gets their own hat ($\frac{1}{5}$), and then multiplying this by the probability of a derangement with $n=4$ (each of the
    remaining four graduates does not get their own cap):
    \begin{align*}
        P(X = 1) &= 5 \cdot \frac{1}{5} \cdot P(\text{derangement for } n = 4 \text{ students}) \\
        &= 1 - \left[ 4 \cdot \frac{1}{4} - \binom{4}{2}\frac{1}{4} \cdot \frac{1}{3} + 
        \binom{4}{3}\frac{1}{4} \cdot \frac{1}{3} \cdot \frac{1}{2} - \binom{4}{4}\frac{1}{4} \cdot \frac{1}{3} \cdot \frac{1}{2} \frac{1}{1}\right] \\
        &= \frac{3}{8} = 0.375
    \end{align*}

    We can see that even for small $n$ (in our case, $n=5$), the expected probabilities approach $\frac{1}{e} \approx 0.36788$.

\end{enumerate}
